```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(edgeR)
  library(sva)
  library(SingleCellExperiment)
  library(TBSignatureProfiler)
  library(tidyverse)
  library(stringi)
  library(hypeR)
  library(DelayedArray)
  library(knitr)
  library(kableExtra)
  library(readxl)
})
  
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(dev = "svg", dpi = 300)
#  library(reshape2)
#  library(RColorBrewer)
#  library(limma)
#  library(knitr)
#  library(pander)
#  library(xtable)
#})

```


# Loading and processing data
```{r load data, include=FALSE}

#indata is a SingleCellExperiment object
indata <- readRDS("new_indata_SCE.rds")

df <- read_xlsx("subject_coldata_cc_excel.xlsx")
df <- column_to_rownames(df, var="...1")


## Removing Outliers
indata <- indata[,-which(colnames(indata) %in% c("10200459B","10200335A","10200332A"))]



# change ages
#colData(indata)$new_age <- df$age

# Want to have 5% present rate
indata <- indata[apply(assay(indata,"counts") != 0, 1, mean)>.2,] 


## Correct some errors in metadata
colData(indata)$Tb_status[grep("074", colnames(indata))] <- "LTBI"
colData(indata)$Tb_status[grep("082", colnames(indata))] <- "LTBI"
colData(indata)$Tb_status[grep("111", colnames(indata))] <- "LTBI"





## Make CPM and log_cpm
#indata  = mkAssay(indata, log = TRUE, counts_to_CPM = TRUE)
#assays(indata)

## Make Severely Malnourished column data
colData(indata)$Tb_status <- factor(gsub("activeTB", "TB", as.character(colData(indata)$Tb_status)))
colData(indata)$bmi_cat2 <- factor(gsub("severely malnourished", "mal", gsub("well nourished", "well", colData(indata)$bmi_cat2)))
colData(indata)$status <- factor(paste(colData(indata)$Tb_status, colData(indata)$bmi_cat2, sep="_"))
table(colData(indata)$status)


colData(indata)$bmi_cat2[grep("082", colnames(indata))] <- factor("mal")
colData(indata)$status[grep("082", colnames(indata))] <- factor("LTBI_mal")

colData(indata)$bmi_cat2[grep("111", colnames(indata))] <- factor("mal")
colData(indata)$status[grep("111", colnames(indata))] <- factor("LTBI_mal")


```

### ComBat Batch correction
```{r corrected data, message=FALSE, results='hide', include = FALSE}

## Make batch variable
batch=rep(1,ncol(indata))
batch[colnames(indata) %in% c("10200247B")]=1
batch[colnames(indata) %in% c("10200227B","10200346A","10200381A","10200308A","10200361A","10200365B","10200467B","10200341A","10200374A", "10200548B","10200343A","10200374B","10200346A","10200377A") ]=1
batch[colnames(indata)%in% c("10200070A", "10200128A","10200056A", "10200075A","10200136A", "10200105A","10200034A","10200058A", "10200007A", "10200064A","10200159A","10200077A","10200073A","10200130A","10200181A","10200059A","10200040B","10200018B","10200065B","10200003B","10200024B","10200004B","10200009B","10200001B","10200057B","10200029B","10200077B","10200035B","10200061B","10200030B","10200063B","10200034B","10200010A","10200288A","10200082A","10200074A","10200060A","10200023A","10200140A","10200111A","10200112A","10200016A","10200015A","10200087A")]=2
colData(indata)$batch <- factor(batch)
rm(batch)

modcombat <- model.matrix(~colData(indata)$bmi_cat2+colData(indata)$Tb_status)

combined_India.combatSeqCorrect <- ComBat_seq(assay(indata, "counts"),
                                      colData(indata)$batch, group=NULL,
                                      covar_mod=modcombat)

assay(indata, "combatseq") <- combined_India.combatSeqCorrect
assays(indata)
indata <- mkAssay(indata, input_name = "combatseq", log = TRUE)


rm(modcombat, combined_India.combatSeqCorrect)

```




### Keep only LTBI
```{r filter, include=FALSE}
indata = indata[,colData(indata)$Tb_status=="LTBI"]
dim(indata)
table(colData(indata)$bmi_cat2)
 
```



### Keep only protein coding genes
```{r filter-by-gene-type, include=FALSE}

#indata <- as(indata, "SingleCellExperiment")

prot_code <- indata@rowRanges@elementMetadata
prot_ <- prot_code[which(prot_code$V2 == "protein_coding"),]
prot <- as.character.Array(prot_@listData$V1)

indata = indata[indata@rowRanges@partitioning@NAMES %in% prot,]

rm(prot_, prot_code, prot)
```


# Distribution of data

## Boxplot of age/bmi_cat2
```{r, include=FALSE, eval=FALSE}

#library(bit64)
df$age <- as.numeric(df$age)
age <- as.numeric(df$age)
names(age) <- (df$SUBJID)


# change to match correct ages to colData
test <- as.data.frame(colData(indata)$SUBJID)
test$ages <- age[match(test$`colData(indata)$SUBJID`, names(age))]

colData(indata)$age <- test$ages

df <- as.data.frame(colData(indata))

boxplot(as.numeric(df$age)~df$bmi_cat2)
```




# SignatureProfiler {.tabset}

## TBSigProfiler Pathways {.tabset}
```{r, message=FALSE, results='hide'}
### make a character list of SUBJID from low BMI to Hi BMI

df <- as.data.frame(indata@colData@listData)
order_bmi <- select(df, SUBJID, BMI)
ordered_bmi <- order_bmi[order(order_bmi$BMI),]

## Keep some pathways in the template as comparison

TBsignatures$Leong_RISK_29 <- c("SRBD1", "ZNF419", "SH2D1B","CTSA", "GSTA4", "AGAP9", "MOB3C", "WARS1", "LUC7L", "ZNRF1", "CIRBP", "PRSS53", "APOL6", "TCN2", "MDN1", "SNRNP70", "SLC3A1", "NAGA", "SPDYE5",  "SPSB1", "CCDC14",  "IL31RA", "DERA", "FUT4", "NEIL1", "ENO3", "CCDC78", "HM13", "ZNF202")

TBsignatures$Zak_RISK_16 <- c("ANKRD22","APOL1","BATF2" ,"ETV7","FCGR1A", "FCGR1B",  "GBP1","GBP2","GBP4","GBP5", "SCARF1","SEPTIN4", "SERPING1", "STAT1","TAP1","TRAFD1")

## Add new signatures to the profiler

TBsignatures$Palmer_40 <- c("IL7", "IL7RA", "IL15", "IL15RA", "BCL2", "BCLXL", "CREBBP", "EP300", "FYN", "IL2RG", "ITGA2B", "JAK1", "JAK3", "LCK", "NMI", "PIK3CA", "PIK3R1", "PTK2B", "STAT5A", "STAT5B", "PDK", "NFKB", "SOCS", "GSK3", "AKT", "MTOR", "CDC25", "PIM1", "PIM2", "P27KIP", "CDC25", "IKK", "IKB", "EIF4E", "BAD", "CDK2", "CDK4", "CDK6", "CD8", "GFI1")

TBsignatures$Naldi_OD_28 <- c("SYK", "RAC1", "TJP1", "CTTN", "PIK3", "ACTB", "RHOA", "GRB2", "SHC1", "SHC2", "SHC3", "SHC4", "SRC", "ARF6", "PIP5K1A", "PIP5K1B", "PIP5K1C", "EZR", "PTEN", "PLCG1", "PLCG2", "PIP4K2A", "PIP4K2B", "GAB2", "SOS1", "PRKCQ", "LCK", "PDPK1")

# Shorten
samp_tbsignatures <- TBsignatures[c("Leong_RISK_29", "Zak_RISK_16", "Palmer_40", "Naldi_OD_28")]


gsva_res <- runTBsigProfiler(indata, useAssay = "log_cpm", algorithm = "GSVA",
                             signatures = samp_tbsignatures, parallel.sz = 4)
ssgsea_res <- runTBsigProfiler(indata, useAssay = "log_cpm", algorithm = "ssGSEA",
                               signatures = samp_tbsignatures, parallel.sz = 4)

plage_res <- runTBsigProfiler(indata, useAssay = "log_cpm", algorithm = "PLAGE", signatures = samp_tbsignatures, parallel.sz = 4)

```

### GSVA {.tabset}

#### Heatmap

```{r subgsva_a_TBsigs}
signatureHeatmap(gsva_res, name="GSVA", signatureColNames = names(samp_tbsignatures),
                 annotationColNames = c("bmi_cat2"),
                 showColumnNames = FALSE,
                 split_heatmap='none')
```


#### Boxplot

```{r boxgsva_TBsigs}
signatureBoxplot(gsva_res, name="GSVA", signatureColNames = names(samp_tbsignatures),
                 annotationColName = c("bmi_cat2"))# , rotateLabels = TRUE)
```

#### Boxplots Single {.tabset}

```{r , results="asis"}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  print(signatureBoxplot(gsva_res,
                         name=i, 
                         signatureColNames = i,
                         annotationColName = c("bmi_cat2"),   
                         rotateLabels = T))

  cat("\n\n")
}

```

#### Signature plots {.tabset}
```{r genes_gsva_TBsigs, results="asis"}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  signatureGeneHeatmap(gsva_res, useAssay="log_cpm",
                     samp_tbsignatures[[i]],
                     name = i, signatureColNames = NULL,
                     annotationColNames = c("bmi_cat2",i),
                     showColumnNames = TRUE, 
                     column_order = ordered_bmi[,1])

  cat("\n\n")
}

```

#### AUC Table {.tabset}
```{r, message = FALSE}
set.seed(0)
tableAUC(gsva_res,
         annotationColName = "bmi_cat2",
         signatureColNames = names(samp_tbsignatures),
         num.boot = 100,
         pb.show = FALSE)
```

#### AUC Boxplots
```{r, message = FALSE}
set.seed(0)
compareBoxplots(gsva_res, annotationColName = "bmi_cat2",
                signatureColNames = names(samp_tbsignatures),
                pb.show = FALSE, fill.col = "blue",
                rotateLabels = TRUE)
```

#### ROC plots
```{r, message = FALSE, fig.height = 9, fig.width = 12}
signatureROCplot_CI(inputData = gsva_res,
                   signatureColNames = names(samp_tbsignatures),
                   annotationColName = "bmi_cat2")

```

#### Separate ROC plots  {.tabset}

```{r, results = 'asis', message = FALSE}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  print(signatureROCplot_CI(inputData = gsva_res,
                   signatureColNames = i,
                   annotationColName = "bmi_cat2",
                   name = paste("ROC plot,", i, sep = " ")))

  cat("\n\n")
}
```


### ssGSEA {.tabset}

#### Heatmap

```{r subssgsea_a_TBsigs}
signatureHeatmap(ssgsea_res, name="ssGSEA", signatureColNames = names(samp_tbsignatures),
                 annotationColNames = c("bmi_cat2"),
                 showColumnNames = FALSE, scale = TRUE,
                 split_heatmap='none')
```

#### Boxplot

```{r boxssgsea_TBsigs}
signatureBoxplot(ssgsea_res, name="ssGSEA", signatureColNames = names(samp_tbsignatures),
                 annotationColName = c("bmi_cat2"), scale = TRUE) #rotateLabels = TRUE,
```

#### Boxplots Single {.tabset}

```{r , results="asis"}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  print(signatureBoxplot(ssgsea_res, name=i, signatureColNames = i,
                 annotationColName = c("bmi_cat2"), rotateLabels = T))

  cat("\n\n")
}

```

#### Signature plots {.tabset}
```{r genes_ssgsea_TBsigs, results="asis"}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  signatureGeneHeatmap(ssgsea_res, useAssay="log_cpm",
                     samp_tbsignatures[[i]],
                     name = i, signatureColNames = NULL,
                     annotationColNames = c("bmi_cat2",i),
                     showColumnNames = FALSE, 
                     column_order =  NULL)

  cat("\n\n")
}

```


#### AUC Table
```{r, message = FALSE}
set.seed(0)
tableAUC(ssgsea_res,
         annotationColName = "bmi_cat2",
         signatureColNames = names(samp_tbsignatures),
         num.boot = 100,
         pb.show = FALSE)
```

#### AUC Boxplots {.tabset}
```{r ssGSEA-AUC-boxplot, message = FALSE}
set.seed(0)
compareBoxplots(ssgsea_res, annotationColName = "bmi_cat2",
                signatureColNames = names(samp_tbsignatures),
                pb.show = FALSE, fill.col = "blue",
                rotateLabels = TRUE)
```

#### ROC plots
```{r, message = FALSE, fig.height = 9, fig.width = 12}
signatureROCplot_CI(inputData = ssgsea_res,
                   signatureColNames = names(samp_tbsignatures),
                   annotationColName = "bmi_cat2")

```

#### Separate ROC plots  {.tabset}

```{r, results = 'asis', message = FALSE}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  print(signatureROCplot_CI(inputData = ssgsea_res,
                   signatureColNames = i,
                   annotationColName = "bmi_cat2",
                   name = paste("ROC plot,", i, sep = " ")))

  cat("\n\n")
}
```


### PLAGE {.tabset}


#### Heatmap

```{r plage_TBsigs}
signatureHeatmap(plage_res, name="PLAGE", signatureColNames = names(samp_tbsignatures),
                 annotationColNames = c("bmi_cat2"),
                 showColumnNames = FALSE,
                 split_heatmap='none')
```


#### Boxplot

```{r boxplage_TBsigs}
signatureBoxplot(plage_res, name="PLAGE", signatureColNames = names(samp_tbsignatures),
                 annotationColName = c("bmi_cat2"))# , rotateLabels = TRUE)
```

#### Boxplots Single {.tabset}

```{r , results="asis"}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  print(signatureBoxplot(plage_res,
                         name=i, 
                         signatureColNames = i,
                         annotationColName = c("bmi_cat2"),   
                         ##violinPlot = TRUE,
                         rotateLabels = T))

  cat("\n\n")
}

```

#### Signature plots {.tabset}
```{r genes_plage_TBsigs, results="asis"}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  signatureGeneHeatmap(plage_res, useAssay="log_cpm",
                     samp_tbsignatures[[i]],
                     name = i, signatureColNames = NULL,
                     annotationColNames = c("bmi_cat2",i),
                     showColumnNames = TRUE, 
                     column_order = ordered_bmi[,1])

  cat("\n\n")
}

```

#### AUC Table {.tabset}
```{r, message = FALSE}
set.seed(0)
tableAUC(plage_res,
         annotationColName = "bmi_cat2",
         signatureColNames = names(samp_tbsignatures),
         num.boot = 100,
         pb.show = FALSE)
```

#### AUC Boxplots
```{r, message = FALSE}
set.seed(0)
compareBoxplots(plage_res, annotationColName = "bmi_cat2",
                signatureColNames = names(samp_tbsignatures),
                pb.show = FALSE, fill.col = "blue",
                rotateLabels = TRUE)
```

#### ROC plots
```{r, message = FALSE, fig.height = 9, fig.width = 12}
signatureROCplot_CI(inputData = plage_res,
                   signatureColNames = names(samp_tbsignatures),
                   annotationColName = "bmi_cat2")

```

#### Separate ROC plots  {.tabset}

```{r, results = 'asis', message = FALSE}
for (i in names(samp_tbsignatures)){

  cat("#####", i, "\n")

  print(signatureROCplot_CI(inputData = plage_res,
                   signatureColNames = i,
                   annotationColName = "bmi_cat2",
                   name = paste("ROC plot,", i, sep = " ")))

  cat("\n\n")
}
```




# Session Info
```{r session info}
sessionInfo()
```